---
title: "Day 8 Computational Session"
format: html
editor: visual
---

## Review basics of ggplot

```{r}
library(tidyverse)
library(here)
library(janitor)


library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
```

Read in data

```{r}
lizards <- read_csv(here("data_tidy","lizards.csv"))
```

Data wrangling to make a new subset from lizards

-   limit data to only contain western whiptails , colorado checkered, eastern fence lizards

-   add a new column with total length converted to centimeters,

-   group by species and sites

-   return a summary table with mean total length and standard deviation of total length by species and site

```{r}
lizards_subset <- lizards %>% 
  filter(common_name %in% c("western whiptail", "colorado checkered","eastern fence")) %>% 
  mutate(total_length_cm = total_length / 10) %>% 
  group_by(common_name, site) %>% 
  summarize(mean_total_length = mean(total_length , na.rm = TRUE),
            sd_total_length = sd(total_length , na.rm = TRUE))
```

```{r}
#how to make a scatterplot total_length vs. weight
ggplot(data = lizards, aes(x = total_length, y = weight))+
  geom_point(aes(color = weight)) +
  scale_y_continuous(limits = c(0,100),
                     breaks = seq(from = 0, to = 100, by = 20 ),
                     expand = c(0,0)) +
  scale_color_gradientn(colors = c("yellow","orange","purple"))
  
  
#scales = free puts each common_name on its own scale.
```

```{r}
#use dplyr::count to find the total number of lizards by common name in the lizards dataset
lizard_counts <- lizards %>% 
  dplyr::count(common_name)
```

```{r}
ggplot(data = lizard_counts , aes(x = n , y = fct_reorder(common_name , n))) +
  geom_col() 
```
